---
title: "Leaflet Exercise"
author: "Nikita Jensen"
date: "Created by Adela Sobotkova, updated by Nikita `r format(Sys.time(), '%d %B %Y')`"
output: 
html_document:
  toc: true
  toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GETTING STARTED WITH LEAFLET

Try to work through down this script, observing what happens in the plotting pane.

Review favorite backgrounds in:
https://leaflet-extras.github.io/leaflet-providers/preview/
beware that some need extra options specified

## Install packages
To install Leaflet package, run this command at your R prompt:
I have CommandOut the packages so I could knit a RMarkdown.
```{r}
#install.packages("leaflet")
```

## We will also need this widget to make pretty maps:
```{r}
#install.packages("htmlwidget")
```

## Activate the libraries
```{r}
library(leaflet)
library(htmlwidgets)
library(tidyverse)
```

# MY TASK NUMBER ONE

# Task 1: Create a Danish equivalent of AUSmap with esri layers, but call it DANmap

## Libraries
```{r}
library(leaflet)
library(htmlwidgets)
library(tidyverse)
```

## Creating labels for my points
```{r}
popup2 = c( "Copenhagen", "Odense", "Aarhus")
```

## Creating a Leaflet Map
```{r}
leaflet() %>%                                 # create a map widget by calling the library
  addProviderTiles("Esri.WorldPhysical") %>%   # add Esri World Physical map tiles
  addAwesomeMarkers(lng = c(12.57, 10.4, 10.21),      # add layers, specified with longitude for 3 points
                    lat = c(55.68, 55.4, 56.16),      # and latitude for 3 points
                    popup = popup2)            # specify labels, which will appear if you click on the point in the map
```

## Looking at Aarhus with setview
```{r}
leaflet() %>%
  addTiles() %>%                              # add default OpenStreetMap map tiles
  addProviderTiles("Esri.WorldImagery",       # add custom Esri World Physical map tiles
                   options = providerTileOptions(opacity=0.5)) %>%     # make the Esri tile transparent
  setView(lng = 10.21076, lat = 56.15674, zoom = 10)              # set the location of the map 
```


## Adding layers
```{r}
leaflet() %>% 
  addTiles() %>% 
  setView( lng = 10.21, lat = 56.16, zoom = 5 ) %>%  # let's use setView to navigate to our area
  addProviderTiles("Esri.WorldPhysical", group = "Physical") %>% 
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>% 
  addProviderTiles("MtbMap", group = "Geo") %>% 
  
  addLayersControl(                                 # we are adding layers control to the maps
    baseGroups = c("Geo","Aerial", "Physical"),
    options = layersControlOptions(collapsed = T))

```

# Check area
```{r}
leaflet() %>% 
  setView(10.21076, 56.15674, zoom = 13) %>%
  addTiles()  # checking I am in the right area
```

## Base location
```{r}
l_dk <- leaflet() %>%   # assign the base location to an object
  setView(10.21076, 56.15674, zoom = 13)
```

## Choosing Esri background layers
```{r}
esri <- grep("^Esri", providers, value = TRUE)

for (provider in esri) {
  l_dk <- l_dk %>% addProviderTiles(provider, group = provider)
}
```

# Creating the map
```{r}
DANmap <- l_dk %>%
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")

DANmap
```

## ADD DATA TO LEAFLET

In this section you will manually create machine-readable spatial data from GoogleMaps: 

### First, go to https://bit.ly/CreateCoordinates1
Enter the coordinates of your favorite leisure places in Denmark 
extracting them from the URL in googlemaps, adding name and type of monument.
Remember to copy the coordinates as a string, as just two decimal numbers separated by comma. 

*Caveats: Do NOT edit the grey columns! They populate automatically!*

### Second, read the sheet into R. You will need gmail login information. 
watch the console, it may ask you to authenticate or put in the number that corresponds to the account you wish to use.

## Libraries
```{r}
library(tidyverse)
library(googlesheets4)
library(leaflet)
```

## Read in a Googlesheet
*If et dosent work use gs4_deauth() *
```{r}
gs4_deauth() 

places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
                     range = "DigitalMethods",
                     col_types = "cccnncnc")
glimpse(places)
```

## Load the coordinates in the map and check:
```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description)

leaflet() %>%
  addTiles() %>%                              # add default OpenStreetMap map tiles
  addProviderTiles("Esri.WorldImagery",       # add custom Esri World Physical map tiles
                   options = providerTileOptions(opacity=0.5)) %>%     # make the Esri tile transparent
  setView(lng = 10.21076, lat = 56.15674, zoom = 10)              # set the location of the map 
```

# Task 2: Read in the googlesheet data you and your colleagues populated with data into the DANmap object you created in Task 1.

## Libraries
```{r}
library(tidyverse)
library(googlesheets4)
library(leaflet)
```

## Read in a Google sheet
*If et dosent work use gs4_deauth()* 
```{r}
places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
                     range = "DigitalMethods",
                     col_types = "cccnncnc")
glimpse(places)
```

## Load the coordinates in the map and check: are any points missing? Why?
```{r}
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description,
             clusterOptions = markerClusterOptions())

leaflet() %>%
  addTiles() %>%                              # add default OpenStreetMap map tiles
  addProviderTiles("Esri.WorldImagery",       # add custom Esri World Physical map tiles
                   options = providerTileOptions(opacity=0.5)) %>%     # make the Esri tile transparent
  setView(lng = 10.21076, lat = 56.15674, zoom = 10)              # set the location of the map 
```

# Task 3: Can you cluster the points in Leaflet? Google "clustering options in Leaflet"

Yes. See solution below.

```{r}
DANmap %>% 
  addMarkers(lng = places$Longitude, 
             lat = places$Latitude,
             popup = places$Description,
             clusterOptions = markerClusterOptions())
```

# Task 4: Look at the map and consider what it is good for and what not.

Maps created in RStudio are a valuable tool for visualizing and analyzing data, offering interactivity, customization, and the ability to convey complex information in a visually engaging way. Like in this map. The map gives a good overview of danish cities, in my example especially Aarhus. The clusterOptions in RStudio's Leaflet package is a valuable tool for customizing the appearance and behavior of marker clusters in your interactive maps. It is particularly useful for creating visually appealing maps with the ability to fine-tune cluster behavior.

# Task 5: Find out how to display notes and classifications in the map.

### Getting rid of NA
```{r}
places <- na.omit(places)
```

## Classification and notes
```{r}
DANmap %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = places$Longitude, 
                   lat = places$Latitude,
                   weight = 2,
                   color = "blue",
                   radius = places$Stars1_5*2,
                   label = places$Placename,
                   popup = paste("Name:", places$Placename, 
                                 "<br>Coordinates:", places$Longitude,",",places$Latitude,
                                 "<br>Type:", places$Type,
                                 "<br>Description:", places$Description,
                                 "<br>Stars:", places$Stars1_5))
```
